---
title: "Twitter US Airline Sentiment"
author: "Carmen Nicolas Baixauli"
subtitle: Analyze how travelers in February 2015 expressed their feelings on Twitter
output: html_document
---


```{r , include=FALSE, cache=FALSE, warning=FALSE}
library(plyr)
library(dplyr)
library(ggplot2)
#library(car)
library(gridExtra)
#library(caTools)
library(tidyr)
library(Hmisc)
library(corrplot)
#library(MASS)
```
### 1. First let's load the data set and take a look at some main features.
```{r , warning=FALSE, echo=FALSE}
file_input = read.csv("/Users/carmennicolasbaixauli/Documents/IE_MBD/3rd Term/RUNNING INNOVATION & CREATION PROCESSES/airline-twitter-sentiment/Tweets.csv")
glimpse(file_input)
#describe(file_input)
```

If there is a variable with too many NA's, it is not going to be useful. The following shows the amount of NA's per variable.
```{r , warning=FALSE, echo=FALSE}
# Here I have places NA in the empty cells
file_input_ = as.data.frame(apply(file_input, 2, function(x) gsub("^$|^ $", NA, x)))
# Checking how many NA's there are per column
apply(file_input_, 2, function(x) sum(is.na(x)))
```
Variables with high ratio of NA's are not going to be representative (for instance tweet_ coord or airline_sentiment _gold)


### 2. Analysis of each variable independently. Some relevant information.
#### 2.1. Number of observations
```{r , warning=FALSE, echo=FALSE}
Twitter<-file_input
nobs<-nrow(Twitter)
cat("Total number of observations:", nobs)
```
#### 2.2. Number of users
```{r , message=FALSE, warning=FALSE, echo=FALSE}
nUsers<-as.numeric(count(distinct(Twitter, name)))#Number of different users
cat("Total number of different users: ", nUsers)
```
#### 2.3. Do all the tweets have a sentiment associated?
```{r , warning=FALSE, echo=FALSE}
#Do all the tweets have a sentiment associated?
noAirline_sentiment<-Twitter[is.na(Twitter$airline_sentiment)]
cat("Total number of different tweets without airline_sentiment detected:", dim(noAirline_sentiment)[2])
```
So yes, all the tweets have a sentiment value associated.

#### 2.4. Negative reasons
If the sentiment is negative, there is a reason that we have obtained this as a result. These are the negative reasons and the proportion of tweets of each of them:
```{r , warning=FALSE, echo=FALSE}
#Negative reasons normal
negatrazon<-Twitter %>%
  group_by(negativereason) %>%
  summarise(Total_complaints=n() )
negatrazon
#Isolate the negtive reason column and count only using table
negat_ra <- Twitter$negativereason
tablanegat_ra <- table(negat_ra)
colors = c("palegreen", "paleturquoise", "palevioletred1", "peachpuff", "pink1", "khaki1", "lightpink", "seagreen3", "plum1", "skyblue2")
pie(tablanegat_ra[2:11], col=colors)
```

The following barchart shows the proportion of these reasons that belong to each airline.

```{r , echo=FALSE}
negat_ra_air <- file_input_[,c(4,6)]
negat_ra_air <- na.omit(negat_ra_air)
qplot(negat_ra_air$negativereason, fill=negat_ra_air$airline, ylab = "Number of Tweets", xlab = "Negative Reason") + theme(axis.title.x = element_text(face="bold", colour="#990000", size=13),
           axis.text.x  = element_text(angle=45, vjust=0.5, size=12))+ theme(axis.title.y = element_text(face="bold", colour="#990000", size=15))+ scale_fill_manual(values=c("deepskyblue", "maroon", "orange", "seagreen3", "midnightblue", "red3"), name  =" ")
```

#### 2.5. Tweet location
```{r , echo=FALSE}
#Tweet_location not useful
cat("The variable tweet_location is not very precise since it is determined by the user. Some examples of why it is not useful are: ", as.character(Twitter$tweet_location[3]),",", as.character(Twitter$tweet_location[11]),",", as.character(Twitter$tweet_location[19]),",", as.character(Twitter$tweet_location[39]), ",",as.character(Twitter$tweet_location[115]))
#Delete location field
Twitter$tweet_location <- NULL
```

#### 2.6. Airlines 
Number of airlines involved in the study and amount of tweets per airline.
```{r, echo=FALSE}
#Number of different airlines and number of tweets per airline
nAirlines<-as.numeric(count(distinct(Twitter, airline)))#Number of different airlines
cat("Total number of different airlines:", nAirlines)
Airlines<-Twitter %>%
  group_by(airline) %>%
  summarise(Total_number_of_Tweets=n() )
Airlines <- as.data.frame(Airlines)
Airlines

plot(Airlines, pch=19,col = "hotpink4", xaxt="n", yaxt="n", ylab="Total number of tweets", xlab="Airline", cex=2.5)
axis(1, at=seq(1, 6, by=1), labels = as.character(Airlines[1:6,1]))
ticks<-seq(400,5000,400)
axis(2, at=ticks, labels=ticks)
par(new=TRUE)
title(main="Number of tweets per airline", col.main="hotpink4", font.main=3, cex.main=1.5)
grid(col = "hotpink4", lwd = 1.5)
```

#### 2.7. Timezones
```{r , echo=FALSE}
#Number of different timezones
nTimezones<-as.numeric(count(distinct(Twitter, user_timezone)))#Number of different user_timezone
cat("Total number of different timezones:", nTimezones, "\n Let's see the number of Tweets per timezone and how many of them do not have any timezone associated (first row)")
Timezones<-Twitter %>%
  group_by(user_timezone) %>%
  summarise(Total_number_of_Tweets=n() )

Timezones<-as.data.frame(Timezones)
Timezones<-as.data.frame(Timezones[order(-Timezones$Total_number_of_Tweets),])
(Timezones)[1:15,]
```

#### 2.8. Sentiment confidence
Distribution of the values of the variable airline_ sentiment_confidence 
```{r , echo=FALSE}
#Analysis of sentiment confidence value
descr_sentconfi <- describe(Twitter$airline_sentiment_confidence)
cat("Five lowest values: ",descr_sentconfi$values[1:5], "\n Five highest values: ",descr_sentconfi$values[6:10], "\n Number of missing values: ",descr_sentconfi$counts[2], "\n Mean value: ",descr_sentconfi$counts[5], "\n Quartiles (0.25, 0.50, 0.75, 1): ",descr_sentconfi$counts[8],descr_sentconfi$counts[9],descr_sentconfi$counts[10],descr_sentconfi$counts[11], "\n Only about 35% of the tweets have a airline sentiment confidence below 0.8. More than 50% of them have the maximum airline sentiment confidence.")

hist(Twitter$airline_sentiment_confidence, col="lightgreen", breaks=25, main ="Airline sentiment confidence", xlab="Confidence")
```

For the tweets with airline sentiment confidence between 0.50 and 0.75 we can find a pattern more or less similar to a normal distribution

```{r , echo=FALSE}
hist(Twitter$airline_sentiment_confidence, col="lightgreen", breaks=250, main ="Airline sentiment confidence", xlab="Confidence", xlim=c(0.55,0.75), ylim=c(0,300))
```

Knowing this information, the airline sentiment variable can be estimated for those tweets where that value is missing. First we calculate the probability of being 1 or less. If we assign a value lower than 1 we can estimate the value using a normal distribution.


#### 2.9. Timestamp (date and time)
Let's take a look at the format of the variable tweet_created and if all the tweets have one assigned.
```{r , echo=FALSE}
#Analyze the tweet_created variable
cat("Do all the tweets have a timestamp associated?", fill=TRUE)
NOtweet_created<-Twitter[is.na(Twitter$tweet_created),]
cat("Total number of different tweets without timestamp detected:", dim(NOtweet_created)[2], "\n This is an example of the variable format: 2015-02-24 11:35:52 -0800" )

```

#### 2.10. Retweet count
```{r , echo=FALSE}
#Retweet frequency
ret_count<-Twitter %>%
  group_by(number_of_retweets=retweet_count) %>%
  summarise(number_of_tweets_retweeted=n() )
ret_count <- as.data.frame(ret_count)
cat("Most of the tweets have 0 retweets. There are ", ret_count[1,2], " tweets not retweeted. Let's visualize the frequency of the tweets that have been retweeted at least once.")
mp <- barplot(ret_count[2:18,2], col = "pink", xlab = "Times retweeted", ylab = "Frequency")
axis(1,at=mp,labels=ret_count[2:18,1])
```

#### 2.11. Gold members
```{r , echo=FALSE}
Twitter_gold <- rbind(Twitter[Twitter$airline_sentiment_gold =="neutral",], Twitter[Twitter$airline_sentiment_gold =="positive",], Twitter[Twitter$airline_sentiment_gold == "negative",])

cat("Total number of gold customers:", dim(Twitter_gold)[1])
```

### 3. Variable transformation

#### 3.1. Split the variable tweet_created
As shown before, the date and the time is not in a easy format to deal with. First we will split it into four variables: year, month, day and time. After that the variable time (which is given in hours, minutes and seconds) is going to be divided into four specific chunks: Morning, Afternoon, Evening and Night.
```{r , echo=FALSE}
#Split date&timestamp in 4 columns: year, month, day and time
Twitter<-Twitter %>%
  separate(tweet_created, c("date", "time"), " ")
Twitter<-Twitter %>%
  separate(date, c("year", "month", "day"), "-")

#Dividing the time variable into specific chunks of time

# [00:00:00, 05:00:00) Night 
Twitter$time<-replace(Twitter$time, "05:00:00" > Twitter$time ,"Night")
# [05:00:00, 12:00:00) Night 
Twitter$time<-replace(Twitter$time, "12:00:00" > Twitter$time ,"Morning")
# [12:00:00, 17:00:00) Night 
Twitter$time<-replace(Twitter$time, "17:00:00" > Twitter$time ,"Afternoon")
# [17:00:00, 21:00:00) Night 
Twitter$time<-replace(Twitter$time, "21:00:00" > Twitter$time ,"Evening")
# [21:00:00, 23:59:59] Night 
Twitter$time<-replace(Twitter$time, "23:59:59" > Twitter$time ,"Night")
```
#### 3.2. Variable selection
Some variables such as twitter_ id, name, text or tweet_coord are not really useful for this analysis so they are going to be removed from the data frame. Since all the tweets are from February 2015, we can also remove the variables month and year. 
```{r , echo=FALSE}
#We are not going to use:twitter_id(1) name(8), text(11), tweet_coord(12), year and month(13, 14)
Twitter_red <- Twitter[c(2,3,4, 5, 6, 7, 9, 10, 15, 16, 17)]
```
#### 3.3. From categorical to numerical
It is easier to work with numerical variables than with categorical. With numerical variables we can measure things such as correlation that can not be calculated with categorical values. These replacements are:

- **Airline sentiment**: _"positive"=1, "neutral"=0, "negative"=-1_

- **Timeslot**: _"Morning"=1, "Afternoon"= 2,"Evening"=3, "Night"=4_

- **Airline**: _"American"=1, "Delta"=2, "Southwest"=3,"United"=4, "US Airways"=5, "Virgin America"=6_

- **Negative reasons**: _"Bad Flight"=1, "Can't Tell"=2, "Cancelled Flight"=3,"Customer Service Issue"=4, "Damaged Luggage"=5, "Flight Attendant Complaints"=6, "Flight Booking Problems"=7, "Late Flight"=8, "longlines"=9, "Lost Luggage"=10_
```{r , echo=FALSE}
#It is easier to work with numeric variables than char. 

#Replacing sentiment values: positive=1, negative= -1, neutral = 0.
Twitter_red_numeric <- Twitter_red
Twitter_red_numeric$airline_sentiment <- revalue(Twitter_red_numeric$airline_sentiment, c("positive"=1, "neutral"=0, "negative"=-1))

#Replacing sentiment gold values: positive=1, negative= -1, neutral = 0.
Twitter_red_numeric$airline_sentiment_gold <- revalue(Twitter_red_numeric$airline_sentiment_gold, c("positive"=1, "neutral"=0, "negative"=-1))

#Replacing timeslot: Morning=1, Afternoon= 2, Evening=3, Night=4.
Twitter_red_numeric$time <- revalue(Twitter_red_numeric$time, c("Morning"=1, "Afternoon"= 2,"Evening"=3, "Night"=4))

#Replacing Airline: "American"=1, "Delta"=2, "Southwest"=3,"United"=4, "US Airways"=5, "Virgin America"=6.
Twitter_red_numeric$airline <- revalue(Twitter_red_numeric$airline, c("American"=1, "Delta"=2, "Southwest"=3,"United"=4, "US Airways"=5, "Virgin America"=6))

#Replacing the 10 different negative reasons: "Bad Flight"=1, "Can't Tell"=2, "Cancelled Flight"=3,"United"=4, "Damaged Luggage"=5, "Flight Attendant Complaints"=6, "Flight Booking Problems"=7, "Late Flight"=8, "longlines"=9, "Lost Luggage"=10.
Twitter_red_numeric$negativereason <- revalue(Twitter_red_numeric$negativereason, c("Bad Flight"=1, "Can't Tell"=2, "Cancelled Flight"=3,"Customer Service Issue"=4, "Damaged Luggage"=5, "Flight Attendant Complaints"=6, "Flight Booking Problems"=7, "Late Flight"=8, "longlines"=9, "Lost Luggage"=10))

```

### 4. Studying relations between variables.

#### 4.1. Tweet count with sentiment (per airline)
Comparison of sentiment type per airline. Each bar represents one of the six airlines. The height of each bar is the total number of tweets per airline. Each color respresents the amount of that tweets that are related to good (green), bad (red) or neutral (blue) experiences.

```{r , echo=FALSE}
qplot(airline, data=Twitter_red, fill=airline_sentiment, ylab = "Number of Tweets", xlab = "Airline", main="Tweet count with sentiment (per airline)") + scale_fill_manual(values=c("indianred2", "lightsteelblue1", "springgreen3"))

```

Are these proportions differents if we only consider gold members?

```{r , echo=FALSE}
Twitter_gold <- rbind(Twitter_red[Twitter_red$airline_sentiment_gold =="neutral",], Twitter_red[Twitter_red$airline_sentiment_gold =="positive",], Twitter_red[Twitter_red$airline_sentiment_gold == "negative",])

Twitter_gold_num <- rbind(Twitter_red_numeric[Twitter_red_numeric$airline_sentiment_gold ==0,], Twitter_red_numeric[Twitter_red_numeric$airline_sentiment_gold ==1,], Twitter_red_numeric[Twitter_red_numeric$airline_sentiment_gold == -1,])

qplot(airline, data=Twitter_gold, fill=airline_sentiment, ylab = "Number of Tweets", xlab = "Airline") + scale_fill_manual(values=c("indianred2", "lightsteelblue1", "springgreen3"))
```

Negative experience percentage is way higher among gold members, the sample of gold customers is not very big (40 observations) so nothing truly insightful can be concluded.


#### 4.2. Linear relation between the main variables
Let's take a look at the linear correlation coeficients between the main variables:
```{r , echo=FALSE}
paracor <- cbind(as.numeric(Twitter_red_numeric$airline_sentiment), as.numeric(Twitter_red_numeric$airline_sentiment_confidence), as.numeric(Twitter_red_numeric$airline), as.numeric(Twitter_red_numeric$retweet_count), as.numeric(Twitter_red_numeric$day), as.numeric(Twitter_red_numeric$time))
colnames(paracor) <- c("airline_sentiment", "airline_sentiment_confidence", "airline", "retweet_count", "day", "time")

MP<-cor(paracor)
colnames(MP)<-colnames(paracor)
col<- colorRampPalette(c("green", "white", "hotpink2"))(15)
corrplot(MP, method = "color", col=col, tl.col="black", tl.srt=45, tl.cex=0.7)
```

The most significant values are between airline sentiment and airline sentiment confidence, nevertheless those values still too close to zero, so that correlations are not relevant.

#### 4.3. Linear relation between the main variables related to bad experiences
The same type of plot is going to be used to check if there is any linear relationship between the bad experiences and the rest of the variables (airline, date, retweets...)

```{r , echo=FALSE}

paracorbad <- na.omit(Twitter_red_numeric)
paracorbad$negativereason <- as.numeric(paracorbad$negativereason)
paracorbad$negativereason <- paracorbad$negativereason-1
paracorbad <- paracorbad[paracorbad$negativereason!=0,]
paracorbad <- paracorbad[,-6]
paracorbad <- paracorbad[,-6]
paracorbad <- paracorbad[,-9]

ParacorBAD<- cbind(as.numeric(paracorbad[,1]), as.numeric(paracorbad[,2]), as.numeric(paracorbad[,3]), as.numeric(paracorbad[,4]), as.numeric(paracorbad[,5]), as.numeric(paracorbad[,6]),as.numeric(paracorbad[,7]), as.numeric(paracorbad[,8])  )
colnames(ParacorBAD) <- colnames(paracorbad)


MPbad<-cor(ParacorBAD[,2:8])
colnames(MPbad)<-colnames(paracorbad[2:8])
col<- colorRampPalette(c("springgreen3", "white", "deepskyblue"))(15)
corrplot(MPbad, method = "color", col=col, tl.col="black", tl.srt=45, tl.cex=0.7)
```

These results are the same as the ones shown by the plot considering the whole dataset. We cannot conclude that there is any linear relationship between variables.

As it was not possible to find any linear relationship between variables, every pair of variables is going to be plotted one against the other to see if we can identify some other kind of relationship (not linear) between them.

#### 4.4. Patterns between variables related to neutral experiences:
The following plot is a graphical representation of the relationship between each pair of variables. The objective is to find non-linear relationships between them. If, for example, the number of retweets decreases exponentially as time (in days) passes, the calculated correlations in the previous chunk would not reflect that relationship. If we detect a notable patter in one of the 15 pairs, we would isolate it and study it further.

```{r , echo=FALSE}
plot(Twitter_red_numeric[,c(1,2,5,8,9,10)], main="Todo")

```

Since we have not detected any special/unique relationhip, we will divide the dataset in 3 parts based on the type of airline sentiment (positive, negative, neutral) and then look for any new non-linear relationships.

#### 4.5. Patterns between variables divided by type of experience:
```{r , echo=FALSE}
par(mfrow=c(1,3))
plot(Twitter_red_numeric[Twitter_red_numeric$airline_sentiment == 1,][,c(2,5,8,9,10)], main="Positive sentiment")
plot(Twitter_red_numeric[Twitter_red_numeric$airline_sentiment == -1,][,c(2,5,8,9,10)], main="Negative sentiment")
plot(Twitter_red_numeric[Twitter_red_numeric$airline_sentiment == 0,][,c(2,5,8,9,10)], main="Neutral sentiment")
```

The graphical representations do not vary even if isolated by sentiment type so we are still unable to find non-linear replationships within the pairs of variables.

### 4. Conclusions

To conclude this EDA we can take note of the following as main points of te story this data tells us.

-There is a total of 14640 observations and 15 variables in the original dataset.

-The % of NA's in airline_ sentiment_ gold, negativereason_gold y tweet _coord is very high (99.7%, 99.7% and 93% respectively) so they will not be useful for our analysis.  

-The 14640 tweets correspond to a total of 7701 unique users

-The most common negative reason was "Customer Service Issue."

-American Airlines is the company with the highest number of tweets classified as having negative sentiment.

-The Tweet location variable is not very reliable as it is determined by the Twitter user and is sometimes a false location (por ejemplo: Lets Play , 1/1 loner squad , Somewhere celebrating life., i'm creating a monster, Earth).

-All the airlines have a tweet count between 222-2913 except Virgin America which only has 504

-The data tells us that there are 86 different time zones, which is obviously not true. This is because some users would put Madrid instead of CEST which contains other cities/countries. However, 11703 of the tweers in our dataset correspond to Eastern/Central/Pacific Time (US & Canada).

-More than 50% of the values of the airline sentiment confidence variable are 1. This means that airline sentiment is very precise in the majority of cases, being greater then or equal to 0.6923 for 75% of the data.

-Airline sentiment shows a normal distribution for those tweets that have a value between 0.55 and 0.75

-the majority of tweets (94.7%) have never been retweeted. THe frequency of those that have been retweeted decreases exponentially.

-The variables associated with gold members have been eliminated since there are only 40 rows where the corresponding values are not null. 

-Intending to calculate correlations and facilitate other types of algorithms, the categorical variables have been transformed to numeric with the following code:

 * Airline sentiment: positive=1, neutral=0, negative=-1

 * Timeslot: "Morning"=1, "Afternoon"= 2,"Evening"=3, "Night"=4

 * Airline: "American"=1, "Delta"=2, "Southwest"=3,"United"=4, "US Airways"=5, "Virgin America"=6

  * Negative reasons: "Bad Flight"=1, "Can't Tell"=2, "Cancelled         Flight"=3,"Customer Service Issue"=4, "Damaged Luggage"=5, "Flight Attendant Complaints"=6, "Flight Booking Problems"=7, "Late Flight"=8, "longlines"=9, "Lost Luggage"=10

-Graphically representing the ratio of positive, negative, and neutral sentiment per airline, we can observe that the majority of tweers are related to a negative sendiment in all 6 cases. Considering the 40 registered gold members, these negative percentages are even more defined, but we can't gather very insightful conclusions since we only have 40 entries for this type of data.

-The linear correlation is not significant for any pair of variables. We cannot find a linear relationship with any of the fields. 

-We have also studied the possibility of detecting nonlinear correlations using graphic representations for each pair of variables, but is has also not been possible.